---
title: "README"
author: "R Ju"
date: "12/9/2020"
output:  
  github_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
library(tidyverse)
library(apex)
library(ape)
library(ggtree)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(geiger)

#read in data
syms <- read.csv("data/syms.csv")
syms_prop <- read.csv("data/syms_prop.csv")
phy1 <- read.tree("output/phy1")
geosyms <- read.csv("data/geosyms.csv")
```


# Phylogenetic signal in stony coral (*Scleractinia*) associations with algal symbionts (*Symbiodinium*)


## Introduction

The goal of my project is to answer the questions...

- Does Symbiodinium community composition recapitulate host phylogeny in Scleractinian corals?
    - Does the level of recapitulation depend on scale/subclade?
    - Does the level of recapitulation depend on the population/biogeography?

- If not, what factors explain the incongruences?
    - How are associations between corals and algae formed?
    - How plastic are such associations?

- How can phylosymbiotic analyses inform related research?
    - What can this phylogenetic analysis tell us about coral-algal plasticity and fitness?

- Do the phylogenetic trees for Scleractinia and Symbiodinium show co-phylogeny?

## Methods

Publicly available data for *Symbiodinium* clade associations with *Scleractinia* species were collected from the [Coral Traits Database](coraltraits.org) along with location metadata. Species with at least 10 *Symbiodinium* clade observations were used in phylogenetic  analysis. Geographic distributions of common *Symbiodinium* clades were visualized using the R packages ___

Gene sequence data for *Scleractinia* species were downloaded from the National Center for Biotechnology Information (NCBI) database. Mitochondrial sequence data for the cytochrome oxidase 1 (CO1) subunit were collected using the NCBI Basic Local Alignment Search Tool (BLAST). A bait sequence for the CO1 gene was chosen within the study taxa (*Acropora digitifera*). CO1 sequence data was also collected for an outgroup species, *Xestospongia testudinaria* (barrel sponge). All sequences were aligned in BLAST before downloading as a FASTA file. Raw alignment sequence data was processed to assign species headers, and concatenated using the R package 'apex' to remove duplicate species. 

Phylogenetic inference was conducted using IQ-TREE (version 1.6.12 ___) on the Yale Center for Research Computing (YCRC) cluster. Maximum likelihood analysis was conducted with 1000 bootstrap replicates. The R package 'ape' was used to root the tree to the specific outgroup species and to trim the tree to exclude tips with no *Symbiodinium* clade data. The tree was visualized as a cladogram with trait data using the R package 'ggtree'.

The character trait of *Symbiodinium* clade associations was coded in three ways:

1. Using only the most-associated clade as a single categorical trait.

2. Treating associations with each clade as six separate binary traits.

3. Consider the number of clades associated as a single categorical trait.

For each method, phylogenetic signal was assessed using the R package 'geiger' to calculate Pagel's lambda.

See final_project.rmd file for analysis code.


## Results


### *Symbiodinium* clade association data

At least 10 observations were found for 100 *Scleractinia* species, with the number of total observations ranging from 10 to 158. Clade C was most commonly observed, followed by clades D, B, and A. Clades F and G were only observed 4 and 2 times, respectively. The number of species in which the clades were observed followed the same trend (Table 1).


| Clade | # observations | # host species |
| ----------- | ----------- | -------- |
| A | 53 | 14 | 
| B | 183 | 21 |
| C | 1846 | 97 |
| D | 388 | 62 |
| F | 4 | 3 |
| G | 2 | 2 |

**Table 1** Table showing *Symbiodinium* clades, the number of observed associations, and the number of species associated with.

Of the 100 study species, 27 were only found to associate with one *Symbiodinium* clade, while 56 were found to associate with 2 clades. 8 and 9 species associated with 3 and 4 clades, respectively. The pie charts in Figure 1 visualize the proportional distribution of *Symbiodinium* clade associations per species. 

```{r fig1, echo = FALSE, fig.width=20, fig.height=20}
syms_prop %>%
  ggplot(aes(x = "", y = prop, fill = value), size = 12) +
  geom_bar(stat = "identity", width = 0.5, color = "white") +
  coord_polar("y", start=0) +
  scale_fill_manual(breaks = c("C", "B", "A",  "D", "F", "G"), values = c("tomato", "cornflowerblue", "seagreen2", "gold", "lightpink", "burlywood"), name = "Symbiodinium clade") +
  facet_wrap(~specie_name) +
  theme_void() + theme(legend.position = "bottom")
```
**Figure 1.** Pie charts showing the proportion of observations attributed to each *Symbiodinium* clade per *Scleractinia* species.

### Phylogenetic inference

```{r fig2, echo = FALSE, fig.height=30, fig.width=20}
plot(phy1) 
nodelabels(phy1$node.label)
```
**Figure 2** Rooted phylogenetic tree for *Scleractinia* with ML support values, generated with CO1 mitochondrial gene sequences. Only tips with *Symbiodinium* clade association data are shown.

Figure 2 shows the rooted phylogenetic tree for all *Scleractinia* species with trait data, generated from IQ-TREE with maximum likelihood support values. The best model fit by Bayesian information criteria (BIC) was TVM+F+I+G4, with a log likelihood value of -15093.3637.

```{r, fig.height = 40, fig.width=40}
tips <- data.frame(tip.label=phy1$tip.label, specie_name=str_replace(phy1$tip.label, "_", " "))

clades <- full_join(tips, filter(syms_prop, specie_name %in% str_replace(phy1$tip.label, "_", " ") == TRUE), by="specie_name") %>%
  filter(is.na(value) == FALSE) %>%
  group_by(tip.label, specie_name) %>%
  dplyr::summarise(value = value) %>%
  filter(is.na(tip.label) == FALSE) %>% #removing internal nodes
  pivot_wider(names_from = value)

ggtree(phy1, branch.length = "none") %<+% clades +
  geom_tiplab(fontface = "italic", size = 10, offset = 0.2) +
  geom_tippoint(aes(color=G), size = 8, position = position_dodge(width = 0)) +
  geom_tippoint(aes(color=F), size = 8) +
  geom_tippoint(aes(color=B), size = 8, position = position_dodge(width = 0.75)) +
  geom_tippoint(aes(color=A), size = 8, position = position_dodge(width = 1.5)) +
  geom_tippoint(aes(color=D), size = 8, position = position_dodge(width = 2.25)) +
  geom_tippoint(aes(color=C), size = 8, position = position_dodge(width = 3)) +
  scale_color_manual(breaks = c("C", "B", "A",  "D", "F", "G"), values = c("tomato", "cornflowerblue", "seagreen2", "gold", "lightpink", "burlywood"), name = "Symbiodinium clade") + 
  xlim(-2, 35) +  
  theme(legend.position = "bottom")
```
**Figure 3** Cladogram of 86 *Scleractinia* species. Colored points at tips correspond to observed *Symbiodinium* clade associations. 

Figure 3 shows the phylogenetic reconstruction as a cladogram, as well as known *Symbiodinium* clade associations. Some grouping in associations with less common clades (A, B) are apparent, for example in *Orbicella* species. However, no clade is grouped monophyletically (excluding G, observed in one species). 

### Phylogenetic signal assessment

To determine if there is strong phylogenetic signal in *Symbiodinium* clade associations in *Scleractinia*, Pagel's lambda was calculated for three different character codings (Table 2). A lambda value of 0 indicates no phylogenetic signal, while a lambda value of 1 indicates high phylogenetic signal (Brownian motion model). 

Method | Coding | lambda value |
| ----------- | ----------- | ---- |
1 | Most-associated clade | 0.733 | 
2a | Clade A association | 0.906 |
2b | Clade B association | 0.994 |
2c | Clade C association | 1.000 |
2d | Clade D association | 0.414 |
3 | # clades associated | 0.399 |
**Table 2** Pagel's lambda values for various character coding methods. 

The first method of character coding by most-associated clade resulted in a lambda value of 0.733, which shows moderate phylogenetic signal. The second method, in which association with each clade is treated as a separate binary trait, yielded >0.9 lambda values for clades A and B, suggesting strong phylogenetic signal. Association with clade C resulted in a 1.000 lambda value — however, as clade C is nearly ubiquitous across the tree due to incomplete sampling, this does not give any information on phylogenetic signal. The lambda value for clade D association was 0.414, indicating a low phylogenetic signal. Similarly, the third method considering the number of clades found in association led to a lamda value of 0.399, also showing low phylogenetic signal.
 
## Discussion

These mixed results indicate that the evolution of symbiotic associations between *Scleractinia* species and *Symbiodinium* clades may be driven by both phylogeny and external factors. 

These results indicate...

The biggest difficulty in implementing these analyses was...

As *Symbiodinium* clade associations are generally characterized as multi-state, categorical traits, many common methods for measuring phylogenetic signal (phylogenetic independent contrasts, phylogenetic least-squares regression, etc.) are not possible.

If I did these analyses again, I would...

## References
